---
title: 'Region preparation'
author: "Melanie (1028-02098-001-001)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    depth: 3
    theme: sandstone
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{1028-02098-001-001}
- \fancyfoot[LE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options:
  chunk_output_type: console
---

# Introduction

The goal of this .Rmd is to prepare and consolidate all peaks for deep learning training and downstream analysis. We will need to assess peak reproducibility, process reproducible peaks, and assess quality metrics of final samples and their respective regions for downstream work.

# Computational Setup

```{r, warning=F, message=F}
#Standard packages
library(rtracklayer) ; library(GenomicRanges); library(magrittr) ; library(Biostrings)
library(ggplot2) ; library(reshape2); library(plyranges); library(Rsamtools); library(parallel)
library(dplyr); library(data.table); library(patchwork); library(readr); library(testit)

#KNITR Options
setwd("/n/projects/mw2098/collaboration/for_carlos/20260205_manuscript/2_modeling/")
figure_filepath<-'figures/2_prepare_peaks'
options(knitr.figure_dir=figure_filepath)

#Lab sources
source("/n/projects/mw2098/shared_code/rscripts/granges_common.r")
source("/n/projects/mw2098/shared_code/rscripts/metapeak_common.r")
source("/n/projects/mw2098/shared_code/rscripts/knitr_common.r")
source("/n/projects/mw2098/shared_code/rscripts/caching.r")
source("/n/projects/mw2098/shared_code/custom/metapeak_functions.R")

#Specific sources
library(BSgenome.Dmelanogaster.UCSC.dm6)
library(TxDb.Dmelanogaster.UCSC.dm6.ensGene)

#Options and configurations
set.seed(seed = 10)
output_seqlen<-1000
dir.create('narrowpeak', showWarnings = F)
dir.create('bed', showWarnings = F)

#Custom variables
bsgenome<-BSgenome.Dmelanogaster.UCSC.dm6
txdb<-TxDb.Dmelanogaster.UCSC.dm6.ensGene
blacklist.gr<-rtracklayer::import('/n/projects/mw2098/genomes/dm6/dm6.blacklist.bed') #obtained from (https://www.nature.com/articles/s41598-019-45839-z)
chroms<-c('chr2L', 'chr2R', 'chr3L', 'chr3R', 'chr4', 'chrX')
```

# Import blacklisted regions

# Define peaks

```{r}
# system('wc -l idr/*')
```

```{r}
peaks.list<-list(

  S2_AntHOX6a_PBX_Flag_nexus = 'idr/S2_AntHOX6a_PBX_Flag_nexus_1_vs_2_idr.txt',
  S2_AntHOX6a_PBX_HA_nexus = 'idr/S2_AntHOX6a_PBX_HA_nexus_1_vs_2_idr.txt',
  
  S2_AntHOX1a_PBX_Flag_nexus = 'idr/S2_AntHOX1a_PBX_Flag_nexus_1_vs_2_idr.txt',
  S2_AntHOX1a_PBX_HA_nexus = 'idr/S2_AntHOX1a_PBX_HA_nexus_1_vs_2_idr.txt',

  S2_Dfd_Hth_HA_nexus = 'idr/S2_Dfd_Hth_HA_nexus_1_vs_2_idr.txt',
  S2_Dfd_Hth_antiExd_nexus = '../1_processing/peaks/S2_Dfd_Hth_anti_EXD_nexus_1_peaks.narrowPeak' #no extra replicate

)

```

# Filter .narrowpeak that you will be using

Here, we will be:

+ removing peaks on chromosome boundaries
+ removing peaks with underlying sequences containing Ns
+ removing peaks from chrY and chrM
+ converting to .narrowPeak format while centering the coordinates around the 'summit' column (col 9).

```{r}
peaks.gr.list<-lapply(names(peaks.list), function(x){
  message(x)
  if(grepl('.txt', peaks.list[[x]])){
    df<-readr::read_tsv(peaks.list[[x]], col_names = c('chrom','start','end','name','score','strand',
                                                       'signalValue', 'pValue', 'qValue', 'peak',
                                                       'localIDR', 'globalIDR',
                                                       'rep1start','rep1end','rep1signal','rep1summit',
                                                       'rep2start','rep2end','rep2signal','rep2summit'))
  } else{
    df<-readr::read_tsv(peaks.list[[x]], col_names = c('chrom','start','end','name','score','strand',
                                                       'signalValue', 'pValue', 'qValue', 'peak'))
  }
  
  gr<-df %>%
    makeGRangesFromDataFrame(keep.extra.columns = T, starts.in.df.are.0based = T) %>%
    plyranges::filter(!overlapsAny(., blacklist.gr, ignore.strand = T)) %>% #remove blacklisted regions
    check_chromosome_boundaries(., 10000, bsgenome) %>%
    plyranges::select(-seqlength_filler, -end_distance, -start_distance) %>%
    keepStandardChromosomes(x = ., pruning.mode = 'coarse') %>%
    plyranges::filter(!(seqnames %in% c('chrY', 'chrM'))) %>%
    
    plyranges::mutate(start = start + peak) %>%
    GenomicRanges::resize(., width = 1, fix = 'start') %>%
    GenomicRanges::resize(., width = output_seqlen, fix = 'center') %>%
    
    plyranges::mutate(peak = floor(output_seqlen/2),  #convert to narrowpeak, broadpeak doesn't have the summit feature
                      seq = getSeq(bsgenome, ., as.character=T)) %>%  
    plyranges::arrange(desc(pValue)) %>% 
    plyranges::filter(!grepl('N', seq)) %>% #Filter out all regions with N's in them
    plyranges::select(-seq)
  
  #Write GRanges
  plyranges::write_narrowpeaks(x = gr, file = paste0('narrowpeak/', tolower(x), '.narrowPeak'))
  rtracklayer::export(object = gr, con = paste0('narrowpeak/', tolower(x), '.bed'))
  return(gr)
})
names(peaks.gr.list)<-names(peaks.list)
```

# Quantify overlaps between peaks

```{r}
#First, mark whether peak set was reproducible
reprod.df<-data.frame(
  from_idr = lapply(peaks.list, function(x) grepl('.txt', x)) %>% unlist(),
  experiment = names(peaks.list)
)
rownames(reprod.df)<-NULL

#Next, do cross comparisons
comparisons.df<-reprod.df$experiment %>%
  tidyr::crossing(e1 = ., e2 = .) %>%
  dplyr::filter(!(e1>=e2))

comparisons_w_values.df<-cbind(comparisons.df,
                               lapply(1:nrow(comparisons.df), function(x){
                                 e1<-peaks.gr.list[[comparisons.df$e1[x]]]
                                 e2<-peaks.gr.list[[comparisons.df$e2[x]]]
                                 
                                 data.frame(e1_n = length(e1),
                                            e2_n = length(e2),
                                            e1_ov_e2 = sum(overlapsAny(e1, e2, ignore.strand = T)),
                                            e2_ov_e1 = sum(overlapsAny(e2, e1, ignore.strand = T)))
                               }) %>% rbindlist()
) %>% dplyr::mutate(e1_ov_rate = round(e1_ov_e2/e1_n, 2), e2_ov_rate = round(e2_ov_e1/e2_n, 2)) %>%
  dplyr::left_join(reprod.df %>% dplyr::rename(e1 = experiment, e1_from_idr = from_idr)) %>%
  dplyr::left_join(reprod.df %>% dplyr::rename(e2 = experiment, e2_from_idr = from_idr)) %>%
  dplyr::mutate(idr_comparison_state = ifelse((e1_from_idr) & (e2_from_idr), 'reproducibe', 'non-reproducible')) %>%
  dplyr::filter(idr_comparison_state != 'non-reproducible')
```

Visualize overlap rates.

```{r}
overall.plot<-ggplot(comparisons_w_values.df, aes(x = e1, y = e2, fill = e1_ov_e2))+
  geom_tile(linewidth = 1, color = 'black')+
  scale_x_discrete(expand = c(0,0)) + scale_y_discrete(expand = c(0,0))+
  ggtitle('Overall peak overlaps')+
  scale_fill_gradientn(colours = viridis(5))+ theme_classic() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

rate_e1.plot<-ggplot(comparisons_w_values.df, aes(x = e1, y = e2, fill = e1_ov_rate))+
  geom_tile(linewidth = 1, color = 'black')+
  scale_x_discrete(expand = c(0,0)) + scale_y_discrete(expand = c(0,0))+
  ggtitle('Overlap rates of x-axis')+
  scale_fill_gradient(high = 'navy', low = 'white', limits = c(0, 1))+ theme_classic() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

rate_e2.plot<-ggplot(comparisons_w_values.df, aes(x = e1, y = e2, fill = e2_ov_rate))+
  geom_tile(linewidth = 1, color = 'black')+
  scale_x_discrete(expand = c(0,0)) + scale_y_discrete(expand = c(0,0))+
  ggtitle('Overlap rates of y-axis')+
  scale_fill_gradient(high = 'darkred', low = 'white', limits = c(0, 1))+ theme_classic() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

plot<-overall.plot + rate_e1.plot + rate_e2.plot + plot_layout(nrow = 1)
ggsave(paste0(figure_filepath, '/reproducible_peak_overlap_rates.png'), height = 6, width = 20)
ggsave(paste0(figure_filepath, '/reproducible_peak_overlap_rates.pdf'), height = 6, width = 20)
```

# Merge all peaks

In both reproducible and non-reproducible formats, merge all peaks for model training.

```{r}
reproducible.gr<-lapply(reprod.df %>% dplyr::filter(from_idr) %>% .$experiment, function(x) peaks.gr.list[[x]]) %>% 
  as('GRangesList') %>% unlist() %>% 
  GenomicRanges::reduce(., ignore.strand = T)

all.gr<-peaks.gr.list %>% 
  as('GRangesList') %>% unlist() %>% 
  GenomicRanges::reduce(., ignore.strand = T)

#Check widths for validity
summary(width(reproducible.gr))
summary(width(all.gr))
```

We can see here that the non-reproducible peaks essentially cover the entire genome and are not reliable. Thus, we will use the reproducible peaks merged together. To consider the entire breadth of the merge, we will tile the merged set. 

```{r}
final.gr<-plyranges::tile_ranges(reproducible.gr, output_seqlen) %>% 
  plyranges::filter(!overlapsAny(., blacklist.gr))
print(length(final.gr))
rtracklayer::export(final.gr, 'bed/all_reproducible_peaks_tiled_expanded.bed')
```

# Session Information

For the purposes of reproducibility, the R/Bioconductor session information is printed below:

```{r sessioninfo}
sessionInfo()
```












