---
title: 'Curation of motifs'
author: "Melanie (1028-02098-001-001)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    depth: 3
    theme: sandstone
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{1028-02098-001-001}
- \fancyfoot[LE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options:
  chunk_output_type: console
---

# Introduction

The goal of this analysis is to:

1. Showcase key motifs found by all binding models.
2. Collect and filter motifs.
3. Merge relevant motifs into sets for downstream analysis.
4. Visualize motif summaries and representations.

# Computational Setup

```{r, warning=F, message=F}
#Standard packages
library(rtracklayer) ; library(GenomicRanges); library(magrittr) ; library(Biostrings)
library(ggplot2) ; library(reshape2); library(plyranges); library(Rsamtools); library(parallel)
library(dplyr); library(data.table); library(patchwork); library(readr); library(testit)

#KNITR Options
setwd("/n/projects/mw2098/collaboration/for_carlos/20260205_manuscript/2_modeling/")
figure_filepath<-'figures/5_curate_motifs'
options(knitr.figure_dir=figure_filepath)

#Lab sources
source("scripts/r/granges_common.r")
source("scripts/r/metapeak_common.r")
source("scripts/r/knitr_common.r")
source("scripts/r/caching.r")
source("scripts/r/metapeak_functions.r")
source("scripts/r/motif_functions.r")

#Specific sources
library(BSgenome.Dmelanogaster.UCSC.dm6)
library(TxDb.Dmelanogaster.UCSC.dm6.ensGene)
library(ggseqlogo)

#Pre-existing variables
threads <- 5
dir.create('bed/mapped_motifs', recursive = T, showWarnings = F)
dir.create('tsv/mapped_motifs', recursive = T, showWarnings = F)
dir.create('rdata/mapped_motifs', recursive = T, showWarnings = F)
bsgenome<-BSgenome.Dmelanogaster.UCSC.dm6
txdb<-TxDb.Dmelanogaster.UCSC.dm6.ensGene

#variable/model information
output_length<-1000
bg_freqs<-c(.27, .21, .21, .27)
```

# Define motifs

Many motifs were found to be contributing but we will focus on the key specific motifs that produce precise TF footprinting. Please reference the `modiscolite` folder for more information and reports on all motifs.

```{r}
motifs.list<-list(
  s2_anthox6a_pbx_flag_nexus_fold1_pbx_counts=list('pattern_1'=c('AntHox6a-PBX', 'pos')),
  s2_anthox6a_pbx_ha_nexus_fold1_pbx_counts=list('pattern_0'=c('AntHox6a-PBX', 'pos')),
  s2_anthox1a_pbx_flag_nexus_fold1_pbx_counts=list('pattern_2'=c('AntHox1a-PBX', 'pos')),
  s2_anthox1a_pbx_ha_nexus_fold1_pbx_counts=list('pattern_5'=c('AntHox1a-PBX', 'neg'))
  
  # s2_dfd_hth_exd_nexus_fold1_dfd_counts = list('pattern_1' = c('AntHox6a-PBX', 'neg')),
  # s2_dfd_hth_ha_nexus_fold1_dfd_counts = list('pattern_0' = c('TTAATGGC', 'pos'))
)
```

# Import hits

Import all hits from the key motifs of interest.

```{r}
hits.list<-lapply(names(motifs.list), function(x){
  message(x)
  info.df<-data.frame(pattern_name = names(motifs.list[[x]]), motif =  motifs.list[[x]][[1]][1], oriented_strand = motifs.list[[x]][[1]][2])
  
  hits.df<-paste0('modiscolite/', x, '/hits.tsv') %>% readr::read_tsv(.)
  hits_filtered.df<-hits.df %>%
    dplyr::mutate(model = x) %>%
    dplyr::filter(pattern_name %in% names(motifs.list[[x]]), metacluster_name == 'pos_patterns') %>%
    dplyr::left_join(info.df) %>%
    as.data.frame
  print(table(hits_filtered.df$motif))
  return(hits_filtered.df)
})
names(hits.list)<-names(motifs.list)
```

# Curate motifs redundancy

Remove the following redundancies:

+ Motifs that are peak-redundant will be mapped across peak "indexes" that are actually located across overlapping genomic regions. 
  + In this case, we are mapping genome-wide, so this step isn't necessary for this exact reason, but we will leave it in because I'm paranoid.
+ Motifs that are palindromic will be mapped on the positive and negative strand with almost identical ranges. 
  + Here, we will remove redundantly mapped motifs and select the unique motif maps that have the highest contribution.

In order to simultaneously manage peak-redundancy and palindromic motifs, we will make miniature islands for each motif group to reduce them.

```{r}
hits_dedup.list<-lapply(names(hits.list), function(modisco_name){
  #Orient strand to match desired aligned convention
  gr<-makeGRangesFromDataFrame(hits.list[[modisco_name]], keep.extra.columns = T, starts.in.df.are.0based = T) %>%
    plyranges::mutate(strand = ifelse(oriented_strand == 'neg',
                                      ifelse(strand == '-', '+', '-'),
                                      strand))

  #Import peaks and reduce them to a deduplicated state to get general peak regions.
  peaks.gr<-gr %>%
    GenomicRanges::reduce(., ignore.strand = T) %>%
    plyranges::mutate(region_id_tmp = 1:length(.))

  #Find overlaps between motifs and peaks
  ov<-findOverlaps(gr, peaks.gr, ignore.strand = T)

  #Assign temporary region ID for deduplication
  gr$region_id_tmp<-NA
  gr$region_id_tmp[ov@from]<-peaks.gr$region_id_tmp[ov@to]

  #Filter any motif that doesn't fit into peaks. These would be weird outliers, shouldn't happen, tbh.
  gr<-gr %>% plyranges::filter(!is.na(gr$region_id_tmp))
  
  #Deduplicate any motifs for palindromic mappings
  dedup.gr<-lapply(unique(gr$motif), function(y){
    print(y)
    motifs.gr<-gr %>% plyranges::filter(motif==y)
    motifs_dedup.gr<-remove_palindromic_motifs_from_bpnet_instances(dfi = motifs.gr, 
                                                                    pattern_to_filter = y, 
                                                                    chrom_name = 'seqnames', start_name = 'start', end_name = 'end', 
                                                                    value_name = 'contrib_match', motif_name = 'motif',
                                                                    region_name = 'region_id_tmp')
    print(paste0(length(motifs_dedup.gr) , ' kept from ', length(motifs.gr)))
    return(motifs_dedup.gr)
  }) %>% as('GRangesList') %>% unlist(.)
  return(dedup.gr)
})
names(hits_dedup.list)<-names(hits.list)
```

Unwrap hits from a list.

```{r}
hits.gr<-hits_dedup.list %>% as('GRangesList') %>% unlist()
names(hits.gr)<-NULL
```

Format the motifs.

```{r}
all_motifs.gr<-hits.gr %>%
  plyranges::mutate(motif_id = 0:(length(.)-1), #0-based indexing later might come in handy
                    seq = getSeq(bsgenome, ., as.character = T))
```

# Export curated motifs

Show final results of motifs returned:

```{r}
table(all_motifs.gr$motif, all_motifs.gr$model)
```

Export curated motifs. In the subsequent .bed file it will be the `[task]_[motif_id]_[region_id]` naming convention.

```{r}
#1-based
motifs.df<-all_motifs.gr %>% as.data.frame()
readr::write_tsv(motifs.df, paste0('tsv/mapped_motifs/all_instances_curated_1based.tsv.gz'))

#0-based
lapply(names(hits_dedup.list), function(x){
  message(x)
  all_motifs.gr %>% plyranges::filter(model==x) %>% plyranges::mutate(name = motif) %>%
    rtracklayer::export(., 
                        paste0('bed/mapped_motifs/all_instances_curated_', x, '_0based.bed'))
})

```

Examine how many motifs show different sequences.

```{r}
all_motifs.gr %>%
  as.data.frame %>%
  dplyr::group_by(model, motif) %>%
  dplyr::slice_max(order_by = contrib_match, n = 250) %>%
  dplyr::summarize(TGATTAAT = sum(grepl('TGATTAAT', seq)),
                   TGATTGAT = sum(grepl('TGATTGAT', seq)),
                   total_motifs = dplyr::n())
```


# Visualize PWMs and CWMs from TF-MoDISco

```{r}
# system(paste0('rm ', figure_filepath, '/unaligned_seqlets/seqlet_logo_*'))

#Plot all seqlet PWM and CWM logos
pwm_and_cwm.logo.list<-lapply(all_motifs.gr$model %>% unique, function(x){
  
  info.df<-data.frame(pattern_name = names(motifs.list[[x]]), motif =  motifs.list[[x]][[1]][1], oriented_strand = motifs.list[[x]][[1]][2])
  
  s.hdf5<-rhdf5::h5read(paste0('modiscolite/', x, '/modisco.h5'),
                        paste0('pos_patterns/', info.df$pattern_name))
  cwm<-s.hdf5$contrib_scores
  pwm<-ppm_to_ic(s.hdf5$sequence, bg_freq = c(.3, .2, .2, .3))
  
  #Orient strand
  if(info.df$oriented_strand=='neg'){
    cwm<-cwm[nrow(cwm):1,ncol(cwm):1]
    pwm<-pwm[nrow(pwm):1,ncol(pwm):1]
  }
  
  rownames(cwm)<-c('A','C','G','T')
  rownames(pwm)<-c('A','C','G','T')
  pwm_and_cwm.list<-list(pwm = pwm, cwm = cwm)
  y_prefix<-paste0(x %>% gsub('_fold1', '', .) %>% gsub('_pbx_counts', '', .) %>% gsub('_dfd_counts', '', .))
  names(pwm_and_cwm.list)<-paste0(y_prefix, c('_pwm', '_cwm'))
  
  pwm_and_cwm.logo<-ggseqlogo(pwm_and_cwm.list, method='custom', seq_type='dna') + 
    facet_wrap(~seq_group, nrow = 1, scales='free') +
    theme(axis.text.x = element_blank())

  return(pwm_and_cwm.logo)
})

pwm_and_cwm_all.logo<-wrap_plots(pwm_and_cwm.logo.list, ncol = 1)
ggsave(paste0(figure_filepath, '/seqlet_logos.png'), pwm_and_cwm_all.logo, height = 8, width = 10)
ggsave(paste0(figure_filepath, '/seqlet_logos.pdf'), pwm_and_cwm_all.logo, height = 8, width = 10)
```

# Visualize ChIP-nexus coverage

Given the set of motifs, what do ChIP-nexus heatmaps look like across the top motifs per logo?

## Allocate coverage and settings

```{r}
nexus_bw.list<-list(
  S2_AntHOX6a_PBX_Flag = list(pos = 'bw/S2_AntHOX6a_PBX_Flag_nexus_combined_positive.bw',
                              neg = 'bw/S2_AntHOX6a_PBX_Flag_nexus_combined_negative.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  S2_AntHOX6a_PBX_HA = list(pos = 'bw/S2_AntHOX6a_PBX_HA_nexus_combined_positive.bw',
                            neg = 'bw/S2_AntHOX6a_PBX_HA_nexus_combined_negative.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  
  S2_AntHOX1a_PBX_Flag = list(pos = 'bw/S2_AntHOX1a_PBX_Flag_nexus_combined_positive.bw',
                              neg = 'bw/S2_AntHOX1a_PBX_Flag_nexus_combined_negative.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  S2_AntHOX1a_PBX_HA = list(pos = 'bw/S2_AntHOX1a_PBX_HA_nexus_combined_positive.bw',
                            neg = 'bw/S2_AntHOX1a_PBX_HA_nexus_combined_negative.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2)
  
  # S2_Dfd_Hth_HA = list(pos = 'bw/S2_Dfd_Hth_HA_nexus_combined_positive.bw',
  #                      neg = 'bw/S2_Dfd_Hth_HA_nexus_combined_negative.bw') %>%
  #   mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  # S2_Dfd_Hth_anti_EXD = list(pos = 'bw/S2_Dfd_Hth_anti_EXD_nexus_combined_positive.bw',
  #                             neg = 'bw/S2_Dfd_Hth_anti_EXD_nexus_combined_negative.bw') %>%
  #   mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2)
  )

#Define how many motifs to consider
top_n = 250
```

## Plot coverage

```{r}
hm_plot_all.list<-lapply(all_motifs.gr$model %>% unique, function(x){
  message(x)
  m.gr<-all_motifs.gr %>% plyranges::filter(model==x) %>% plyranges::arrange(desc(contrib_match)) %>% 
    head(n=top_n) %>% 
    plyranges::arrange(contrib_match)
  print(m.gr$seq %>% table %>% sort(decreasing = T))
  
  plots_across_motif.list<-mclapply(names(nexus_bw.list), function(y){
    nexus.df<-plot_nexus_heatmap(regions.gr = GenomicRanges::resize(m.gr, width = 1, fix = 'center'), 
                                   sample = nexus_bw.list[[y]], upstream = 150, downstream = 151, 
                                   title = y, order = 'asis', removal_threshold = .8, normalize_threshold = .999,
                                   return_only_df = T) #.8, .99
    
    nexus.plot<-ggplot()+
      geom_tile(data=nexus.df %>% dplyr::filter(strand == 'pos'), aes(x=position, y=region, fill=signal))+
      geom_tile(data=nexus.df %>% dplyr::filter(strand == 'neg'), aes(x=position, y=region, fill=-signal), alpha=.5)+
      ggtitle(paste0('Motif: ', m.gr$motif %>% unique, '\nfrom model\n', x), subtitle = paste0('ChIP: ', y))+
      scale_x_continuous("Position (bp)", expand = c(0, 0))+
      scale_y_continuous("Regions", expand = c(0, 0))+
      scale_fill_gradientn(colors = c('#08306b', '#08519c' , '#2171b5', '#4292c6', '#6baed6',
                                      '#9ecae1', '#c6dbef','#deebf7', '#f7fbff', "white", '#fff5f0',
                                      '#fee0d2','#fcbba1','#fc9272','#fb6a4a','#ef3b2c','#cb181d','#a50f15','#67000d'), name = "norm. signal") +
      theme_classic()+
      theme(text=element_text(size=14), legend.position = "none", panel.grid = element_blank(), panel.border = element_blank())
    return(nexus.plot)
  }, mc.cores = 4)
  
  hm.plot<-wrap_plots(plots_across_motif.list, nrow = 1) 
  return(hm.plot)
})

hm_all.plot<-wrap_plots(hm_plot_all.list, ncol = 1) 
ggsave(paste0(figure_filepath, '/nexus_heatmap_across_all_motifs.png'), hm_all.plot, height = 16, width = 16)
ggsave(paste0(figure_filepath, '/nexus_heatmap_across_all_motifs.pdf'), hm_all.plot, height = 16, width = 16)
```

Report underlying sequences.

```{r}
lapply(all_motifs.gr$model %>% unique, function(x){
  message(x)
  m.gr<-all_motifs.gr %>% plyranges::filter(model==x) %>% plyranges::arrange(desc(contrib_match)) %>% 
    head(n=top_n) %>% 
    plyranges::arrange(contrib_match)
  print(m.gr$seq %>% table %>% sort(decreasing = T)) %>% head(n=5)
})
```


## Clarify coverage by sequence

Does Dfd-HA bind in exclusively heterodimeric or monomeric conformations? Sort upstream sequences from the TTAAT core and check.

```{r}
# all_motifs.gr %>% plyranges::filter(model=='s2_dfd_hth_ha_nexus_fold1_dfd_counts') %>% as.data.frame %>% ggplot(., aes(x = contrib_magnitude))+geom_density()
# 
# m.gr<-all_motifs.gr %>% plyranges::filter(model=='s2_dfd_hth_ha_nexus_fold1_dfd_counts') %>% 
#   plyranges::filter(contrib_magnitude>=0.1) %>%
#   plyranges::mutate(upstream_seq = getSeq(bsgenome, 
#                                           GenomicRanges::resize(
#                                             GenomicRanges::resize(
#                                               GenomicRanges::resize(., 1, 'start'), 4, 'end'), 3, 'start'))) %>%
#   plyranges::arrange(upstream_seq, desc(contrib_match))
# 
# 
# seq.list<-plot_sequence(m.gr, title = 'Dfd motifs', genome = bsgenome, window = 15, show = F)
# seq.plot<-seq.list$plot
# 
# nexus.df<-plot_nexus_heatmap(regions.gr = GenomicRanges::resize(m.gr, width = 1, fix = 'center'), 
#                                    sample = nexus_bw.list$S2_Dfd_Hth_HA, upstream = 150, downstream = 151, 
#                                    title = y, order = 'asis', removal_threshold = .8, normalize_threshold = .99,
#                                    return_only_df = T)
#     
# nexus.plot<-ggplot()+
#   geom_tile(data=nexus.df %>% dplyr::filter(strand == 'pos'), aes(x=position, y=region, fill=signal))+
#   geom_tile(data=nexus.df %>% dplyr::filter(strand == 'neg'), aes(x=position, y=region, fill=-signal), alpha=.5)+
#   ggtitle(paste0('Sequence separated motif'))+
#   scale_x_continuous("Position (bp)", expand = c(0, 0))+
#   scale_y_continuous("Regions", expand = c(0, 0))+
#   scale_fill_gradientn(colors = c('#08306b', '#08519c' , '#2171b5', '#4292c6', '#6baed6',
#                                   '#9ecae1', '#c6dbef','#deebf7', '#f7fbff', "white", '#fff5f0',
#                                   '#fee0d2','#fcbba1','#fc9272','#fb6a4a','#ef3b2c','#cb181d','#a50f15','#67000d'), 
#                        name = "norm. signal") +
#   theme_classic()+
#   theme(text=element_text(size=14), legend.position = "none", panel.grid = element_blank(), panel.border = element_blank())
# 
# seq_separated.plot<-seq.plot + nexus.plot
# seq_separated.plot
# ggsave(paste0(figure_filepath, '/nexus_heatmap_across_seq_separated_dfd.png'), seq_separated.plot, height = 6, width = 12)
# ggsave(paste0(figure_filepath, '/nexus_heatmap_across_seq_separated_dfd.pdf'), seq_separated.plot, height = 6, width = 12)
```

Find where the precise upper footprints are.

```{r}
# mp.plot<-plot_nexus_metapeak(GenomicRanges::resize(m.gr, 1, 'center'), nexus_bw.list$S2_Dfd_Hth_HA)
# mp.plot + theme_bw()

```

```{r}
# #Collect positive strand coverage
# m_w_peaks.gr<-m.gr %>% 
#   keepStandardChromosomes(.) %>%
#   plyranges::mutate(monomeric_peak = regionSums(GenomicRanges::resize(., 3, 'center'), nexus_bw.list$S2_Dfd_Hth_HA$pos),
#                     hereodimeric_peak = regionSums(GenomicRanges::resize(GenomicRanges::resize(GenomicRanges::resize(., 1, 'center'), 12, 'end'), 3, 'start'), nexus_bw.list$S2_Dfd_Hth_HA$pos))
# 
# #Contrast positive strand coverage
# 
# cor(m_w_peaks.gr$monomeric_peak, m_w_peaks.gr$hereodimeric_peak)
# 
# ggplot(m_w_peaks.gr %>% as.data.frame, aes(x = monomeric_peak, y = hereodimeric_peak))+
#   geom_point() + 
#   scale_x_continuous(limits = c(0, 200))+
#   scale_y_continuous(limits = c(0, 200))
                    
```

Seems like there is no hidden upstream heterodimeric component for this experiment.

# Visualize predictions across key locus

Here, Carlos chose to showcase the in vivo ChIP-nexus specificity at the following locus: ush (chr2L: 515500-516000)

## Allocate coverage

Given the fold that uses this for the test set in training, we generate data here that predicts this (fold2).

```{r}
pred.bw.list<-list(
  S2_AntHOX6a_PBX_Flag = list(pos = 'preds/s2_anthox6a_pbx_flag_nexus_fold2_pbx_pos_all.bw',
                              neg = 'preds/s2_anthox6a_pbx_flag_nexus_fold2_pbx_neg_all.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  S2_AntHOX6a_PBX_HA = list(pos = 'preds/s2_anthox6a_pbx_ha_nexus_fold2_pbx_pos_all.bw',
                            neg = 'preds/s2_anthox6a_pbx_ha_nexus_fold2_pbx_neg_all.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  
  S2_AntHOX1a_PBX_Flag = list(pos = 'preds/s2_anthox1a_pbx_flag_nexus_fold2_pbx_pos_all.bw',
                              neg = 'preds/s2_anthox1a_pbx_flag_nexus_fold2_pbx_neg_all.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  S2_AntHOX1a_PBX_HA = list(pos = 'preds/s2_anthox1a_pbx_ha_nexus_fold2_pbx_pos_all.bw',
                            neg = 'preds/s2_anthox1a_pbx_ha_nexus_fold2_pbx_neg_all.bw') %>%
    mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2)
  
  # S2_Dfd_Hth_anti_EXD = list(pos = 'preds/s2_dfd_hth_exd_nexus_fold2_dfd_pos_all.bw',
  #                            neg = 'preds/s2_dfd_hth_exd_nexus_fold2_dfd_neg_all.bw') %>%
  #   mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2),
  # S2_Dfd_Hth_HA = list(pos = 'preds/s2_dfd_hth_ha_nexus_fold2_dfd_pos_all.bw',
  #                      neg = 'preds/s2_dfd_hth_ha_nexus_fold2_dfd_neg_all.bw') %>%
  #   mclapply(., function(x) {rtracklayer::import.bw(x, as="RleList")}, mc.cores = 2)
)

locus.list<-list(
  anthox6a = GRanges('chr2L', IRanges(516200, 516600), strand = '*'),
  anthox1a = GRanges('chr2R', IRanges(21027100, 21027200), strand = '*')
  # dfd =  GRanges('chr2L', IRanges(14565500, 14565800), strand = '*')
)

contrib.bw.list<-list(
  S2_AntHOX6a_PBX_Flag = 'shap/s2_anthox6a_pbx_flag_nexus_fold1_pbx_counts.bw' %>% rtracklayer::import.bw(., as="RleList"),
  S2_AntHOX6a_PBX_HA = 'shap/s2_anthox6a_pbx_ha_nexus_fold1_pbx_counts.bw' %>% rtracklayer::import.bw(., as="RleList"),
  
  S2_AntHOX1a_PBX_Flag = 'shap/s2_anthox1a_pbx_flag_nexus_fold1_pbx_counts.bw' %>% rtracklayer::import.bw(., as="RleList"),
  S2_AntHOX1a_PBX_HA = 'shap/s2_anthox1a_pbx_ha_nexus_fold1_pbx_counts.bw' %>% rtracklayer::import.bw(., as="RleList")
)

```

## Visualize predictions

```{r}
lapply(names(locus.list), function(x){
  message(x)
  
  locus.gr<-locus.list[[x]]
  
  #Collect contribution scores
  tracks<-names(contrib.bw.list)[grepl(x, tolower(names(contrib.bw.list)))]
  
  contrib.mat.list<-lapply(tracks, function(y){
    df<-data.frame(
      signal = standard_metapeak(GenomicRanges::resize(locus.gr, width = 1, fix = 'center'), contrib.bw.list[[y]],
                                 upstream = floor(width(locus.gr)/2), downstream = ceiling(width(locus.gr)/2)) %>% .$reads,
      seq = stringr::str_split(string = getSeq(bsgenome, locus.gr, as.character = T), pattern = '') %>% unlist()
    ) %>%
      dplyr::mutate(row = 1:nrow(.)) %>%
      tidyr::pivot_wider(id_cols = row, names_from = seq, values_from = signal) %>%
      dplyr::arrange(row)
    df[is.na(df)]<-0
    
    mat<-df %>% dplyr::select(A, C, G, T) %>% as.matrix
    mat[is.na(mat)]<-0
    mat<-t(mat)
    return(mat)
  })
  names(contrib.mat.list)<-tracks
  
  logos.plot<-ggseqlogo(contrib.mat.list, method='custom', seq_type='dna') +
    theme_logo() + 
    facet_wrap(~seq_group, ncol=1, scales='free_x') +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  ggsave(paste0(figure_filepath, '/featured_locus_logos_', x, '.png'), logos.plot, height = 6, width = 12)
  ggsave(paste0(figure_filepath, '/featured_locus_logos_', x, '.pdf'), logos.plot, height = 6, width = 12)
  
  
  #Collect prediction scores
  tracks<-names(pred.bw.list)[grepl(x, tolower(names(pred.bw.list)))]
  
  preds_all.df<-mclapply(tracks, function(y){
    preds.df<-exo_metapeak(gr = GenomicRanges::resize(locus.gr, width = 1, fix = 'center'), sample = pred.bw.list[[y]], 
                           upstream = floor(width(locus.gr)/2), downstream = ceiling(width(locus.gr)/2)) %>%
      dplyr::mutate(genomic_coordinate = start(GenomicRanges::resize(locus.gr, width = 1, fix = 'center')) + tss_distance,
                    sample = y, type = 'predicted ChIP-nexus')
    return(preds.df)
  }, mc.cores = 2) %>% rbindlist()
  
  obs_all.df<-mclapply(tracks, function(y){
    obs.df<-exo_metapeak(gr = GenomicRanges::resize(locus.gr, width = 1, fix = 'center'), sample = nexus_bw.list[[y]], 
                           upstream = floor(width(locus.gr)/2), downstream = ceiling(width(locus.gr)/2)) %>%
      dplyr::mutate(genomic_coordinate = start(GenomicRanges::resize(locus.gr, width = 1, fix = 'center')) + tss_distance,
                    sample = y, type = 'observed ChIP-nexus')
    return(obs.df)
  }, mc.cores = 2) %>% rbindlist()
  
  all.df<-rbind(preds_all.df, obs_all.df) 
  
  pred.plot<-ggplot(all.df, aes(genomic_coordinate, reads))+
    geom_area(aes(group=interaction(sample,strand), fill = type))+
    ggtitle(x)+
    scale_x_continuous(name = "Position (chr2L, bp)")+
    scale_y_continuous(name = "Reads")+
    scale_fill_manual(values = c('#e08214','#8073ac'), name = "ChIP") +
    facet_wrap(type ~ sample, scale = 'free')+
    theme_classic()+
    theme(text=element_text(size=14))
  pred.plot
  
  ggsave(paste0(figure_filepath, '/featured_locus_predictions_', x, '.png'), pred.plot, height = 6, width = 12)
  ggsave(paste0(figure_filepath, '/featured_locus_predictions_', x, '.pdf'), pred.plot, height = 6, width = 12)
  
})
```

# Conclusions

In conclusion, motifs were curated based on multiple redundancies and saved to 1-based and 0-based coordinates. These will be the reference motifs that we base subsequent analysis on.

# Session Information

For the purposes of reproducibility, the R/Bioconductor session information is printed below:

```{r sessioninfo}
sessionInfo()
```












